---
title: "Data608-Module-01-HomeWork"
author: "Vinayak Kamath"
date: "02/14/2021"
output:
  html_document: 
    toc: true
    toc_depth: 4
    number_sections: true
    theme: united
    highlight: tango
  pdf_document: default
---


```{r message=FALSE, warning=FALSE}
#Loading additional libraries
library(dplyr)
library("ggplot2")     

```

**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}

inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)

```

And lets preview this data:

```{r}

head(inc)

```
```{r}

summary(inc)

```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

1. ***=> Using the `str` function we can see there are 50001 observations in the data set and there are 8 variables:   ***  
    ```{r}
    str(inc)
    ```  

2. ***=> We can see that their are 52 unique values in the variable `State` and this includes Puerto Rico (PR) and Washington D.C. (DC) in addition to the 50 states:  ***  
    ```{r}
    
    str(inc$State)
    ```  

3. ***=> Not all observations have count of employees in it. There are 12 observations with NA values:  ***  
    ```{r}
    summary(inc$Employees)
    filter(inc, is.na(Employees))
    ```
      
4. ***=> The max value fro the variable `Rank` shown is 5000; where as there are 5001 observations; telling the variable `Rank` is not unique:  ***  
    ```{r}
    summary(inc$Rank)
    data1 <- inc %>%
      group_by(Rank) %>%
      summarise(n = n()) %>%
      filter(n != 1)
    
    data1
    # we can see there are 2 records each for the Rank 3424 and 5000
    ```
      
  
---  

## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r}

# Grouping the data to the count by State
data1 <- inc %>%
  group_by(State) %>%
  summarise(n = n())

# Plot the bar chart 
barplot(data1$n,names.arg=data1$State,xlab="Count of Companies",ylab="State",col=data1$State, main="Distribution of Companies by State", horiz = T
        ,  las=2 , cex.names=.3,  space =0.2, font=2 , cex.axis = .7)

  
```

---  

## Question 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r}

# R's `complete.cases()` function to get the observations with full data. 12 records that had NA will be discarded:
complete_data <- filter(inc, complete.cases(inc) )
str(complete_data)

# Grouping  by State to get the state with 3rd most companies in it:
complete_data_3rd_state <- complete_data %>%
  group_by(State) %>%
  summarise(n = n()) %>%
  mutate(ranks = order(order(n, decreasing=T))) %>%
  filter(ranks == 3)

# We can see NY as the state with 311 companies in it:
complete_data_3rd_state

# Getting the data set to plot having total, mean and median of employees grouped by Industry:
data2 <- complete_data %>%  
  filter(State == complete_data_3rd_state$State) 

# Ggplot2 plot that shows the average and/or median employment by industry for companies in this state:
ggplot(data2, aes(x = Industry, y = log(Employees) , fill= Industry)  ) + 
  geom_boxplot(outlier.colour="seagreen", outlier.shape=16, outlier.size=2)    +
  theme(legend.position = "none") +
  coord_flip() +
  labs(title="Employment by Industry for Companies in New York State", y="Employees(log)", x="Industry")

```

---  

## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}

# Grouping the data to the count by industries
data3 <- complete_data %>%
  group_by(Industry) %>%
  summarise(total_employees = sum(Employees), total_revenue = sum(Revenue)) %>%
  mutate(revenue_per_employee = total_revenue/total_employees)

# Check the Data:
data3

# Plot the bar chart 
barplot(data3$revenue_per_employee,names.arg=data3$Industry,col=data3$Industry, main="Distribution of Revenue Per Employee by Industry", horiz = F ,ylab="Revenue Per Employee", xlab="Industry" , las=2 , cex.names=.5,  space =1 , cex.axis = .5 )

 
  
```

---  
